---
title: How to use Well Plate Maker
author:
- name: Hélène Borges
  affiliation: Université Grenoble Alpes
  email: borges.helene.sophie@gmail.com
abstract: >
  WPM is a shiny application deployed in the form of an R package. Its 
  objective is to allow a user to generate a well plate plan in order to 
  perform his experiments by controlling __batch effects__ (in particular 
  preventing plate edge effects). The algorithm for placing the samples is 
  inspired by the backtracking algorithm.
vignette: >
  %\VignetteIndexEntry{How to use Well Plate Maker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document:
    toc: yes
    toc_float: true

---

# Getting started
## Prerequisites
This tutorial explains how to use the Well Plate Maker shiny application.
Make sure you are using the R 3.6.0 version.
Also, check that you have these packages installed (with their dependencies):

* `r CRANpkg("shiny")`
* `r CRANpkg("shinydashboard")`
* `r CRANpkg("shinyWidgets")`
* `r CRANpkg("shinycustomloader")`
* `r CRANpkg("DT")`
* `r CRANpkg("RColorBrewer")`
* `r CRANpkg("logging")`
* `r CRANpkg("dplyr")`
* `r CRANpkg("stringr")`
* `r CRANpkg("ggplot2")`

## How to install
From __GitHub__ (if you have devtools installed):
```{r install from github, eval=FALSE}
> devtools::install_github(repo = "HelBor/wpm", )
```

From __BioConductor__:
```{r install from BioConductor, eval=FALSE}
> if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

> BiocManager::install("wpm")
```


## Launch WPM

Whether you use RStudio or simply work with an R console, the procedure remains the same:

```{r launch wpm, eval=FALSE}
> library(wpm)
> wpm()
```
If everything is in order, a new window will open in your default browser

# How to use WPM

Since WPM is a GUI, the idea is to just provide a minimum of parameters to the application. 
No programming skills are required. WPM supports multiple plates and places samples in a balanced way among the plates.

WPM has 4 main panels: __Home__, __Parameters__, __Results__ and __Help__.

## The __Home__ Panel

This page briefly presents the aim of the app, shows the last package version and explains how to cite us to support our work.

## The __Parameters__ Panel

This page is divided into 6 main steps, detailed below:

### Upload dataset

Here you must provide a __CSV__ file containing at least one kind of information: the sample names.
```{r, echo=FALSE}
knitr::kable(
  data.frame("Sample" = c("s1","s2","s3","s4"))
)
```
If your samples belong to different groups that you want to distinguish, then you must provide the additional column "Group".
```{r, echo=FALSE}
knitr::kable(
  data.frame("Sample" = c("s1","s2","s3","s4"), 
             "Group" = c("A","A","B","C"))
)
```



### Plate(s) dimensions

### Forbidden wells

### Blanks

### Not Randomized wells

### Number of iterations





**2)** Specify the plate dimensions and their number (the user can choose between 6,24,  48,  96,  386,  1534  and  custom)  (WPM  checks  that  all  the  given  settings  arecompatible)

**3)** Specify the __Forbidden well__: These  wells  will  not  be  filled  with  any  kind  of  sample. We simply do not want to fill them (e.g. the coins of the plate), or in case of dirty wells, broken pipettes, etc.

**4)** Specify the __Blanks__: correspond to solution without biological sample in it. Provide the neighborhood constraints, which depend on the "Blank" mode chosen. (Shouldn't samples from the same group be found side by side?)

**5)** Specify the __Not Randomized samples__: correspond to Quality Control samples or standards.

**6)** Choose a maximum number of iterations that WPM can do to find a solution,then start WPM. If the samples do not have a group, then the samples will be placedcompletely randomly on the plates. If there are groups, wpm will use an algorithminspired by the backtracking algorithm (in order to place the samples in the wellswhile respecting the specified constraints.).


## The __Results__ Panel

This Panel allows you to look after the final dataset containing the wells chosen for each sample and a plot of your final well-plate map. Dataframe and plots are downloadable separately.

## The __Help__ Panel


# Citing Our work
> The published article of the project will be linked here.

# SessionInfo
```{r SessionInfo}
sessionInfo()
```




